---
title: "MCBoostSurv - Basics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MCBoostSurv - Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN,
  eval = NOT_CRAN
)
```


```{r setup}
library("mcboost")
library("mlr3")
library("mlr3proba")
library("mlr3pipelines")
library("mlr3learners")
library("tidyverse")
set.seed(27099)
```


## Minimal Example: McBoostSurv

To show the basic functionality of `MCBoostSurv`, we provide a minimal example on 
the standard survival data set rats. After loading and pre-processing the data, we train
a `mlr3learner` on the training data. We instantiate a `MCBoostSurv` instance
with the default parameters. Then, we run the `$multicalibrate()` method on our data to start multi-calibration in survival analysis. With `$predict_probs()`, we can get
multicalibrated predictions. 

```{r}

#prepare task 
task = tsk("rats")
prep_pipe = po("encode", param_vals = list(method="one-hot")) 
prep = prep_pipe$train(list(task))[[1]]

#split data
train = prep$clone()$filter(1:199)
val = prep$clone()$filter(200:250)
test = prep$clone()$filter(256:300)

# get trained survival model 
baseline = lrn("surv.ranger")$train(train)

# initialize mcboost
mc_surv = MCBoostSurv$new(init_predictor = baseline)

# multicalibrate model 
mc_surv$multicalibrate(data = val$data(cols = val$feature_names), 
                       labels = val$data(cols = val$target_names))

# get new predictions
mc_surv$predict_probs(test$data(cols = test$feature_names))

```
## What does mcboost do?

Internally mcboostsurv runs the following procedure `max_iter` times (similar ro `mcboost`, just for distributions over time):

1. Predict on X using the model from the previous iteration, `init_predictor` in the first iteration.
1. Compute the residuals `res = y - y_hat` for all time points
1. Split predictions into `num_buckets` according to `y_hat` and time.
1. Fit the auditor (`auditor_fitter`) (here called`c(x)`) on the data in each bucket with target variable `r`.
1. Compute `misscal = mean(c(x) * res(x))`
1. if `misscal > alpha`:
    For the bucket with highest `misscal`, update the model using the prediction `c(x)`.
    else:
    Stop the procedure



## Multicalibrate model trained on PBC data

Based on this, we can now show multicalibration on a data set with two sensitive attributes (age and gender). Again, we load and pre-process the data.

### Load Dataset
```{r}
library(survival)
data_pbc = pbc %>%
    mutate(status = if_else(status == 2, 1, 0)
    ) %>%
    select(-id) %>%
    drop_na()

task_pbc = TaskSurv$new("pbc", backend = as_data_backend(data_pbc), 
                        time = "time", event = "status")


#Create data split

train_test = rsmp("holdout", ratio = 0.8)$instantiate(task_pbc)
train_g = train_test$train_set(1)
test_ids = train_test$test_set(1)
train_val = rsmp("holdout", ratio = 0.75)$instantiate(task_pbc$clone()$filter(train_g))
train_ids = train_val$train_set(1)
val_ids = train_val$test_set(1)

# Train distributional survival model 

xgb_distr = as_learner(ppl("distrcompositor",
               learner = as_learner(prep_pipe %>>% lrn("surv.xgboost"))))

xgb_distr$train(task_pbc$clone()$filter(train_ids))



```

### Mutlicalibrate survival model with validation data

```{r}
# initialize mcboost
 mcboost_learner = as_learner(
   prep_pipe %>>% ppl_mcboostsurv(
     learner = as_learner(prep_pipe %>>% xgb_distr), 
     param_vals = list(
       alpha = 1e-6,
       eta = 0.2,
       time_buckets = 2,
       num_buckets = 1 )
  )
)

# multicalibrate model
mcboost_learner$train(task_pbc$clone()$filter(val_ids))

# get new predictions
test_task = task_pbc$clone()$filter(test_ids)
pred_pbc_mc = mcboost_learner$predict(task_pbc$clone()$filter(test_ids))

pred_pbc_xgb = xgb_distr$predict(task_pbc$clone()$filter(test_ids))

```

### Development of IBS in the defined subgroups
```{r}

pred_pbc_xgb$score(msr("surv.graf"))
pred_pbc_mc$score(msr("surv.graf"))
```



